Stimjim: open source hardware for precise electrical stimulation

### 2019-05-12 benchmarking/paper/draft2.Rnw
1) cleanup. 2) moved stimjim functions to library. 3) updates to draft

\begin{table}
\small{
\begin{tabular}{|l|c|c|c|c|}
\hline
                                & Stimjim     & PulsePal     & STG400x (Multichannel Systems)    & PHM-15x (Med associates) \\
\hline
 Number of output channels      & 2           & 4            &  2-8                              & 2                        \\
 Current output mode            & X           &              &  X                                & X                        \\
 Voltage output mode            & X           & X            &  X                                &                          \\
 Outputs are electrically isolated           & X           &              &  X                                & X                        \\
 Onboard measurement            & X           &              &                                   & Optional                 \\
 Compliance voltage             & $\pm$ 13.7V & $\pm$ 15V    & $\pm$ 120V                        & $\pm$ 45V                \\
 Fastest pulse                  & 50us        & 10us         & 20us                              & 60us                     \\
 Cost (USD)                     & \$194       & \$200        & \$4131 (2 channels)               &                          \\
                                &             &              & \$7462  (4 channels)              &                          \\
                                &             &              & \$10327 (8 channels)              &                          \\
 Open source                    & X           & X            &                                   &                          \\
 \hline
\end{tabular}
}
\caption{\label{tab:stimulatorComparison} Comparison of Stimjim vs other related hardware. Cost was obtained as either the complete cost of the bill of materials (Stimjim and PulsePal) or quoted prices from manufacturers (STG400x and PHM-15x).}
\end{table}

@@

\section*{Results}

@@

<<benchmarkFigure, echo=FALSE,fig.height=8,fig.width=8,fig.cap="\\label{fig:benchmarking}">>=

par(mar=c(4,5,2,2), mfrow=c(4,3))
#matplot(sapply(res[1:60], function(x) x$pulseMat[,1]), col=c(1,2), type='l', lty=1)
plot(paramMat[1:30,3], sapply(res[seq(1,60,by=2)], function(x) x$IPI.summary[3])*1e6, 
  log='xy', main='median IPI vs target', cex=0.75, xlab='Target IPI', ylab='Median measured IPI')
points(paramMat[1:30,3], sapply(res[seq(2,60,by=2)], function(x) x$IPI.summary[3])*1e6,
  col='red', cex=0.75)
abline(a=0,b=1)

#matplot(dendRites::offset(sapply(res[61:120], function(x) x$pulseMat[,1])), col=c(1,2), type='l', lty=1)
plot(paramMat[1:30,3], sapply(res[seq(1,60,by=2)], function(x) x$IPI.sd)*1e6, 
  log='x', main='IPI sd', cex=0.75, xlab='Target IPI', ylab='Measured IPI std. dev.')
points(paramMat[1:30,3], sapply(res[seq(2,60,by=2)], function(x) x$IPI.sd)*1e6,
  col='red', cex=0.75)

@@

@@

>>>>

\begin{table}
\small{
\begin{tabular}{|l|c|c|c|c|}
\hline
                                & Stimjim     & PulsePal     & STG400x (Multichannel Systems)    & PHM-15x (Med associates) \\
\hline
 Number of output channels      & 2           & 4            &  2-8                              & 2                        \\
 Current output mode            & X           &              &  X                                & X                        \\
 Voltage output mode            & X           & X            &  X                                &                          \\
 Outputs are electrically isolated           & X           &              &  X                                & X                        \\
 Onboard measurement            & X           &              &                                   & Optional                 \\
 Compliance voltage             & $\pm$ 13.7V & $\pm$ 15V    & $\pm$ 120V                        & $\pm$ 45V                \\
 Fastest pulse                  & 50us        & 10us         & 20us                              & 60us                     \\
 Cost (USD)                     & \$194       & \$200        & \$4131 (2 channels)               &  \$6211 (plus \$4438     \\
                                &             &              & \$7462  (4 channels)              &  for necessary software) \\
                                &             &              & \$10327 (8 channels)              &                          \\
 Open source                    & X           & X            &                                   &                          \\
 \hline
\end{tabular}
}
\caption{\label{tab:stimulatorComparison} Comparison of Stimjim vs other related hardware. Cost was obtained as either the complete cost of the bill of materials (Stimjim and PulsePal) or quoted prices from manufacturers (STG400x and PHM-15x).}
\end{table}

@@

\section*{Results}
We developed Stimjim to be an precision stimulus generator and isolator in a single package. Stimjim is based on the Teensy 3.5 microcontroller, which utilizes a 32-bit Arm Cortex-M4 processor running at 120MHz. Each stimulating channel includes a current source based on an improved Howland current pump, and a voltage source (an op-amp), driven by a 16-bit digital-to-analog converter (DAC). The final output of each channel is selected by a 4-way switch, such that either channel can be configured as a current output, voltage output, grounded, or disconnected (high-Z). Toe ensure the output is connected to a load and not either an open or short circuit, each channel also has a analog-to-digital converter able to read either the output voltage or the output current (via a sense resistor in series with the current output).

Our design was made in Kicad, an open-source PCB CAD package. Schematic, layout, bill of materials, and build instructions are included in the supplement and in a git repository (https://bitbucket.org/natecermak/stimjim).

@@

<<benchmarkFigure, echo=FALSE,fig.height=7,fig.width=8,fig.cap="\\label{fig:benchmarking}">>=

ex1 = sapply(dataFiles[seq(15,1,length.out = 6)], function(f){
  dat = matrix(readBin(f, what='integer', n=1e7, size=2, endian = 'big'),ncol=2,byrow=TRUE)
  fp = min(which(dat[,1] > 100))
  dat[fp + (-1e3:3e4),1]*vpau
})

par(mar=c(4,4,1,1), mfrow=c(3,4), mgp=c(2,0.7,0))
matplot(1e3/fs*1:nrow(ex1), dendRites::offset(ex1,15), col=1, type='l', lty=1, bty='n', yaxt='n', ylab='', xlab='time (ms)'); 
plot(paramMat[1:30,3], sapply(res[seq(1,60,by=2)], function(x) x$IPI.summary[3])*1e6, 
  log='xy', main='median IPI vs target', cex=0.75, xlab='Target IPI', ylab='Median measured IPI')
points(paramMat[1:30,3], sapply(res[seq(2,60,by=2)], function(x) x$IPI.summary[3])*1e6,
  col='red', cex=0.75)
abline(a=0,b=1)

plot(paramMat[1:30,3], sapply(res[seq(1,60,by=2)], function(x) x$IPI.sd)*1e6, 
  log='x', main='IPI sd', cex=0.75, xlab='Target IPI', ylab='Measured IPI std. dev.')
points(paramMat[1:30,3], sapply(res[seq(2,60,by=2)], function(x) x$IPI.sd)*1e6,
  col='red', cex=0.75)

@@

ex2 = sapply(dataFiles[seq(52,31,length.out=6)], function(f){
  dat = matrix(readBin(f, what='integer', n=1e7, size=2, endian = 'big'),ncol=2,byrow=TRUE)
  fp = min(which(dat[,1] > 100))
  dat[fp + (-1e3:1e4),1]*vpau
})

matplot(1e3/fs*1:nrow(ex2), dendRites::offset(ex2,15), col=1, type='l', lty=1, bty='n', yaxt='n', ylab='', xlab='time (ms)');

@@

ex3 = sapply(dataFiles[seq(81,61,length.out=6)], function(f){
  dat = matrix(readBin(f, what='integer', n=1e7, size=2, endian = 'big'),ncol=2,byrow=TRUE)
  fp = min(which(dat[,1] > 10))
  dat[fp + (-300:700),1]*vpau
})

matplot(1e3/fs*1:nrow(ex3), dendRites::offset(ex3,15), col=1, type='l', lty=1, bty='n', yaxt='n', ylab='', xlab='time (ms)');
